---
title: "Halloween Mini-Project"
author: "Marcus Lau"
format: gfm
---



##Importing Candy data
```{r}
candy_file <- "candy-data.csv"

candy = read.csv("candy-data.txt", row.names=1)
head(candy)
```

>Q1. How many different candy types are in this dataset?

```{r}
nrow(candy)
```

>Q2. How many fruity candy types are in the dataset?

```{r}
sum(candy$fruity)
```


##Favorite Candies


>Q3. What is your favorite candy in the dataset and what is it’s winpercent value?

```{r}
candy["Twix", ]$winpercent
```

```{r}
candy["3 Musketeers", ]$winpercent
```


>Q4. What is the winpercent value for “Kit Kat”?

```{r}
candy["Kit Kat", ]$winpercent
```


>Q5. What is the winpercent value for “Tootsie Roll Snack Bars”?

```{r}
candy["Tootsie Roll Snack Bars", ]$winpercent
```




```{r}
library("skimr")
skim(candy)
```

```{r}
skimr::skim(candy)
```


>Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?
yes, the winpercent


>Q7. What do you think a zero and one represent for the candy$chocolate column?
Whether its chocolate or not

>Q8. Plot a histogram of winpercent values

```{r}
hist(candy$winpercent)
```


```{r}
library(ggplot2)
ggplot(candy)+
  aes(winpercent)+
  geom_histogram(bins=10, col="yellow", fill="blue")
```

>Q9. Is the distribution of winpercent values symmetrical?
no the distribution is not symmetrical

>Q10. Is the center of the distribution above or below 50%?
below 50%

>Q11. On average is chocolate candy higher or lower ranked than fruit candy?

```{r}

chocolate.inds <- as.logical(candy$chocolate)
chocolate.wins <- candy[chocolate.inds,]$winpercent
chocolate.wins
#Average Chocolate
mean(chocolate.wins)
```

```{r}
fruity.inds <- as.logical(candy$fruity)
fruit.wins <- candy[fruity.inds,]$winpercent
fruit.wins
#Average Fruity
mean(fruit.wins)
```
>On average, chocolate is ranked higher than fruit candy.


>Q12. Is this difference statistically significant?

Yes it is significant since  p-value = 2.871e-08

```{r}
t.test(chocolate.wins, fruit.wins)
```

>Q13. What are the five least liked candy types in this set?

```{r}
head(candy[order(candy$winpercent),], n=5)

```

>Q14. What are the top 5 all time favorite candy types out of this set?

```{r}
tail(candy[order(candy$winpercent),], n=5)

```




##Overall Candy Rankings
>Q15. Make a first barplot of candy ranking based on winpercent values.

```{r}
library(ggplot2)

ggplot(candy) + 
  aes(winpercent, rownames(candy)) +
  geom_col()
```


>Q16. This is quite ugly, use the reorder() function to get the bars sorted by winpercent?

```{r}
library(ggplot2)

ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col()

```


```{r}
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"

ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col(fill=my_cols) 
```

```{r}
ggsave("tmp.png")
```

>Q17. What is the worst ranked chocolate candy?
Sixlets

>Q18. What is the best ranked fruity candy?
Starburst

##Price Percents

>Q19. Which candy type is the highest ranked in terms of winpercent for the least money - i.e. offers the most bang for your buck?
Reeses Minature

```{r}
library(ggrepel)

# How about a plot of price vs win
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text_repel(col=my_cols, size=3.3, max.overlaps = 5)

```


>Q20. What are the top 5 most expensive candy types in the dataset and of these which is the least popular?


```{r}
ord <- order(candy$pricepercent, decreasing = TRUE)
head( candy[ord,c(11,12)], n=5 )
```

5 most expensive candies are Nik L Nip, Ring Pop, Nestle Smarties, Hershey Krackel and Hersheys Milk Chocolate. Least popular is Nik L Nip.



##Exploring Correlation Structure
```{r}
library(corrplot)
```

```{r}
cij <- cor(candy)
corrplot(cij)
```


>Q22. Examining this plot what two variables are anti-correlated (i.e. have minus values)?
2 variables with anti correlation are chocolate and fruity.

>Q23. Similarly, what two variables are most positively correlated?
2 variables most positively correlated is chocolate and winpercent.



##Principal Component Analysis

```{r}
pca <- prcomp(candy, scale=TRUE)
summary(pca)
```

```{r}
pca$rotation[,1]
```

#Comparing PC1 vs PC2

```{r}
plot(pca$x[,1:2], col=my_cols, pch=16)
```

```{r}
# Make a new data-frame with our PCA results and candy data
my_data <- cbind(candy, pca$x[,1:3])

```


```{r}
p <- ggplot(my_data) + 
        aes(x=PC1, y=PC2, 
            size=winpercent/100,  
            text=rownames(my_data),
            label=rownames(my_data)) +
        geom_point(col=my_cols)

p
```

```{r}
library(ggrepel)

p + geom_text_repel(size=3.3, col=my_cols, max.overlaps = 7)  + 
  theme(legend.position = "none") +
  labs(title="Halloween Candy PCA Space",
       subtitle="Colored by type: chocolate bar (dark brown), chocolate other (light brown), fruity (red), other (black)",
       caption="Data from 538")

```

```{r}
#library(plotly)
#ggplotly(p)
```


```{r}
par(mar=c(8,4,2,2))
barplot(pca$rotation[,1], las=2, ylab="PC1 Contribution")
```

>Q24. What original variables are picked up strongly by PC1 in the positive direction? Do these make sense to you?
Fruity,hard, and pluribus are in a positive direction.
